---
title: Regeneron effectivness  matching
output: 
  html_document:
    toc: false
    toc_depth: 3
    toc_float: true
    code_folding: hide
date: "`r format(Sys.time(), '%d %B %Y')`"
params:
  min_age: 12
  input_table: regeneron_effectivness
  omicron_table: covid_sequencing
  create_tables: 1
  create_table1: 1
  create_table3: 1
  num_controls: 5
  filter_missing_covariates: 1
  psm: 0

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(root.dir = "/regeneron_effectiveness/code/R/Results/")
```

## parameters:
  * input_table =  `r params$input_table`\
  * ages =`r params$min_age` - and above\
  * create_tables = `r params$create_tables`\
  * filter_missing_covariates = `r params$filter_missing_covariates`\
  * distance method for matching  (0 = mahalanobis, 1 = PS) =  `r params$psm` \
Note: scroll down to define matching formulas

```{r message=FALSE, warning=FALSE, include=FALSE}
library(plyr)
library(tidyverse)
library(RODBC)
library(dbplyr)
library(gtsummary)
library(ggplot2)
library(MatchIt)
library(glue)
library(feather)
library(cobalt)
library(knitr)
library(cri.utils)
library(lubridate)
library(flextable)
library(formula.tools)
library(dplyr)
```


```{r include=FALSE}
source("./utils.R")
```


```{r include=FALSE}
options(scipen = 999)

set.seed(2023)
# Note: for optimal matching is better to leave numerical varaible uncategorized
# and just recode missing value as an extreemly low number like -999
# so a non missing value will not be matched to a missing value 
NUMERIC_NA = as.numeric(-999)

```

```{r include=FALSE}
con <- cri_connection()
```


```{r read sql,  include=FALSE}
filename = glue("feathers/{params$input_table}.feather")
if (file.exists(filename)) {
  df_raw = read_feather (filename)
} else {
df_raw = connect_table(con, params$input_table) %>%
  distinct() %>%
  collect()
write_feather(df_raw, filename)
}
```

```{r}
df_omc = connect_table(con, params$omicron_table) %>%
  select(-teudat_zehut, -shem_prati, -shem_mishpacha)  %>%
  filter(status %in% c('omicron_high_prob', 'omicron_sequenced')) %>% 
  distinct() %>%
  collect()
```

## num initial regeneron recipients
```{r initial stats, eval=FALSE, include=FALSE}
reg_df %>%
  rename_with ( simplify) %>%
  filter(!is.na(regeneron_date)) %>% 
  select(teudat_zehut_with_bikoret) %>% 
  dplyr::summarise(n=n()) %>% 
  pull(n)
```
```{r echo=FALSE}
df_raw = df_raw %>%
  print_total( "initial entries") %>% 
  filter(invalid_tz == 0) %>% 
  print_total( "after filtering invalid TZ entries") %>% 
  left_join(df_omc, by=c('guid_tz' =  'Guid_tz')) %>% 
  filter(is.na(status) | ( status != 'omicron_sequenced' & status != 'omicron_high_prob')) %>% 
  print_total( "after filtering omicron")
```


```{r initial cleaning, echo=FALSE}
df_raw = df_raw %>%
  filter(sw_continuous_membership == 1) %>%
  print_total("after filtering non continuous members") %>% 
  mutate(psw_nursing_home = replace_na(psw_nursing_home, 0)) %>%
  mutate(age = pn_age) %>%
  mutate(SES = case_when(is.na(SES) ~  as.character('unknwon'), T ~ as.character(SES))) %>%
  mutate(SES = plyr::mapvalues(SES, from = c("1", "2", "3"), to = c('Low', 'Medium', 'High'))) %>% 
  mutate(SES = as.factor(SES)) %>% 
  mutate(SES = fct_relevel(SES, "Low","Medium", "High")) %>%
  mutate(across(contains('days_from_dose'), function(x) { ifelse(x<0, as.numeric(NA_integer_),x)})) %>%
  distinct(guid_tz, index_date, .keep_all = TRUE) %>% 
  print_total("after distinct(guid_tz, index_date)")

```

```{r}
# negative days_from_dose_1 means vaccination after the positive PCR - this is irrlevant for our study
df_raw = df_raw %>%
  mutate(days_from_dose_1 = case_when(days_from_dose_1 <=0 ~ NA_real_, T ~ days_from_dose_1))
```

```{r age filtering, echo=FALSE}
  df = df_raw %>%
    filter(between(age, params$min_age, 120)) %>% 
  print_total("after age filtering")
```


```{r time to outcome calc, echo=FALSE}
  fwp_end_date =  max(as.Date((df_raw$retrieval_date)))
df = df %>%
 mutate (days_to_hosp    = as.numeric(difftime(first_time_hosp, index_date,       units = "days"))) %>%
 mutate (days_to_severe  = as.numeric(difftime(first_time_severe_clean, index_date,units = "days"))) %>%
 mutate (days_to_death   = as.numeric(difftime(date_ptira_corona, index_date,     units = "days"))) %>%
 mutate (days_to_end_fwp = as.numeric(difftime(fwp_end_date, index_date,     units = "days"))) %>%
 filter (is.na(days_to_hosp) | days_to_hosp >= 0) %>%
 print_total("after filtering hosp outcome before index_date ") %>%
 filter (is.na(days_to_severe) | days_to_severe >= 0) %>%
 print_total("after filtering severe outcome before index_date ") %>%
 filter (is.na(days_to_death) | days_to_death >= 0) %>%
 print_total("after filtering death outcome before index_date ") %>%
 mutate (days_from_treatment_to_hosp = as.numeric(difftime(first_time_hosp, regeneron_date,  units = "days"))) %>% 
 filter (is.na(days_from_treatment_to_hosp) | days_from_treatment_to_hosp >= 0) %>%
 print_total("after filtering hospitalization outcome before intervention ") %>% 
 mutate (days_from_treatment_to_severe = as.numeric(difftime(first_time_severe_clean, regeneron_date,  units = "days"))) %>% 
 filter (is.na(days_from_treatment_to_severe) | days_from_treatment_to_severe >= 0) %>%
 print_total("after filtering severe outcome before intervention ") 
```

```{r}
df = df %>%
  mutate(vacc1_ww_raw = round(as.numeric(difftime(Date_hisun_dose_1, ymd('2020-12-20'), units = "week")))) %>% 
  filter(is.na(vacc1_ww_raw) | vacc1_ww_raw >= 0) %>% 
  print_total(glue("after filtering invalid first vaccine dose date ")) 
hist(df$vacc1_ww_raw, breaks =100)
```

```{r num weeks from start of campaign to 1st vaccination, include=FALSE}
df = df %>%
  mutate(vacc1_ww = case_when(is.na(vacc1_ww_raw)                  ~ NUMERIC_NA, 
                                          T                        ~ vacc1_ww_raw)) %>%
  mutate(simplified_vacc1_ww = case_when(is.na(vacc1_ww_raw)       ~ 'unvacc', 
                                          between(vacc1_ww,0,3)    ~ '0-3',
                                          between(vacc1_ww,4,7)    ~ '4-7',
                                          between(vacc1_ww,8,10)   ~ '8-10',
                                          between(vacc1_ww,11,19)  ~ '11-19',
                                          T                        ~ '20+')) %>% 
  mutate(simplified_vacc1_ww = fct_relevel(simplified_vacc1_ww, 'unvacc','0-3', '4-7', '8-10','11-19', '20+'))  
```
```{r is waning immunity, echo=FALSE}
df = df %>% 
  mutate(days_from_last_dose_to_infection = pmin(days_from_dose_1, days_from_dose_2, days_from_dose_3, na.rm = T)) %>% 
  mutate(sw_waned_immunity = case_when(num_vacc_doses >=2 & days_from_last_dose_to_infection < 150    ~ 0, 
                                                                T                                     ~ 1))
```

```{r eval=FALSE, include=FALSE}
df %>% 
  group_by(sw_waned_immunity) %>%
  count()
```



```{r applying complete cases on BMI and smoking}
if (params$filter_missing_covariates ==1) {
df1 = df %>% 
  filter(!is.na(bmi_cat)) %>%
  print_total( "after removing entries with missing BMI") %>% 
  filter(!is.na(pc_smoke_desc)) %>% 
  print_total( "after removing entries with missing smoking status")
} else {
  df1 = df
}
```

```{r params filtering, include=FALSE}
df2 = df1 %>%
  mutate(SES = factor(SES)) %>%
  mutate(exposure = if_else(!is.na(regeneron_date), 1,0)) %>% 
  mutate(exposure_cat = factor(exposure)) %>% 
  mutate(bmi = case_when(is.na(bmi) ~ 0,  
                                   T ~ bmi)) %>%
  mutate(bmi_cat = case_when(is.na(bmi_cat) ~ 'unknown',
                             T ~ bmi_cat)) %>% 

  mutate(pc_smoke_desc = case_when(is.na(pc_smoke_desc) ~ 'unknown',
                                   T ~ pc_smoke_desc)) %>% 
 
  mutate(pc_ses_comb20 = case_when(is.na(pc_ses_comb20) ~ NUMERIC_NA, 
                                  T ~ as.numeric(pc_ses_comb20))) 
  
```

```{r include=FALSE}
df3 = df2 %>% 
  mutate(num_cdc_risk_factors = case_when(pn_cdc_risk_factors >= 4 ~ as.character('4+'),
                                          T                        ~ as.character(pn_cdc_risk_factors))) %>% 
  mutate(sector_cat = case_when(
    sector == 'כללי' ~ 'general',
    sector == 'מגזר החרדים' ~ 'orthodox_jew',
    sector == 'מגזר ערבי' ~ 'arab',
    T ~ 'unknown'
    )) %>%
  mutate(sector_cat = factor(sector_cat)) %>%
  mutate (sector_cat = fct_relevel(sector_cat, "general")) %>%
  mutate(pn_flu_vacc_5yr = if_else(pn_flu_vacc_5yr >=5, as.numeric(5), as.numeric(pn_flu_vacc_5yr))) %>%
  mutate(pn_gp_visits_1yr = if_else(pn_gp_visits_1yr >=5, '5+', as.character(pn_gp_visits_1yr))) %>%
  mutate(age_group = case_when(
    between(age, 12, 18) ~ "12-18",
    between(age, 19, 29) ~ "19-29",
    between(age, 30, 39) ~ "30-39",
    between(age, 40, 49) ~ "40-49",
    between(age, 50, 59) ~ "50-59",
    between(age, 60, 69) ~ "60-69",
    between(age, 70, 74) ~ "70-74",
    age >= 75 ~ "75+",
  )) %>%
  mutate(obesity = between(bmi, 30,40)) %>% 
  mutate(severe_obesity = case_when(bmi >= 40 ~ 1, T ~0)) %>% 
  mutate(overweight = between(bmi, 25,29.999)) %>% 
  mutate(pcr_date = index_date) %>% 
  mutate(sw_heart = pmax(psw_cdc_possible_cereb_vasc, psw_heart)) %>% 
  mutate(sw_diabetes = pmax(psw_dm_type1, psw_dm_type2)) %>% 
  mutate(sw_respiratory = pmax(psw_cdc_possible_asthma, psw_cdc_certain_copd, psw_cdc_possible_other_resp)) 

  
```  



# Table1
#### Initial population stratified by epxosure
```{r table1, echo=FALSE}
if (params$create_tables == 1 | params$create_table1 ==1) {
  table1 = tbl_summary(
      df3 %>% 
select (exposure_cat , age, age_group , sector_cat , sex, SES , pn_flu_vacc_5yr, pn_pcr_tests_9m, bmi_cat,  pc_smoke_desc,  num_cdc_risk_factors, psw_active_malig, psw_cdc_certain_ckd, psw_cdc_certain_copd, psw_cdc_certain_heart,
        severe_obesity,obesity, overweight, psw_preg, psw_cdc_certain_sickle, psw_cdc_certain_dm2,psw_dm_type1, 
 psw_cdc_possible_asthma, psw_cdc_possible_cereb_vasc, psw_cdc_possible_other_resp,psw_cdc_possible_htn, 
 sw_imunosuppressed, psw_cdc_possible_neuro, psw_hepatic, num_vacc_doses, sw_waned_immunity, vacc1_ww_raw, contains('serology')),
      by = exposure_cat, # split table by group
      label = variables_mappings,
      statistic = all_continuous() ~ "{mean} ({sd}), {median} ({p25},{p75})",
      missing = "ifany", # don't list missing data separately
    ) %>%
      modify_header(label = "**Variable**") %>% # update the column header
      modify_header(update = all_stat_cols() ~ "**{ifelse(level==0,'No Regeneron','Regeneron')}**, N = {n} ({style_percent(p, symbol = TRUE)})") %>%
      bold_labels() %>%
      add_overall() 
  table1 %>%
    gtsummary::as_flex_table() %>% 
    save_as_docx(path= glue("Results/table1.docx"))
  
  table1
}

```

```{r echo=FALSE}
filename = glue("feathers/{params$input_table}_cleaned.feather")
write_feather(df3, filename)
```

```{r warning=FALSE, include=FALSE}
if (params$psm == 0) {
formula = exposure_cat ~ 
            age + 
            pc_ses_comb20 + 
            vacc1_ww +
            bmi
} else {
  formula = exposure_cat ~ age_group + bmi_cat + simplified_vacc1_ww+ SES+ sex+ sector_cat+sw_imunosuppressed + psw_preg
}

exact_formula = ~ 
            sex +
            sector_cat + 
            sw_imunosuppressed +
            psw_preg

```
#####################################################
#         M A T C H I N G
#####################################################

```{r matching slow, echo=FALSE, message=TRUE, warning=FALSE}
matched_df = match_multiple_controls_without_immortal_time_bias(df3, formula, exact_formula, params$num_controls , distance_method = case_when(params$psm == 1 ~ "ps", T  ~"mahalanobis"))
```

```{r}
unmatched_df =  
  df3 %>% 
  filter(exposure_cat == 0) %>% 
  left_join(matched_df %>%
              filter(exposure_cat == 0) %>% 
              select(exposure_cat,guid_tz)
            ,by='guid_tz') %>% 
  mutate(is_matched = case_when(is.na(exposure_cat.y) ~ 'No' , T ~ 'Yes' ))
```
```{r eval=FALSE, include=FALSE}
unmatched_df %>% 
  group_by(is_matched) %>% 
  count()
```


### number of matched subjects in each arm
```{r echo=FALSE}
matched_df %>% 
  group_by(arm) %>% 
  count() %>% 
  pull(n)
```

### Overall Matching formula:
`r deparse(formula)`\

### Exact matching variables :
`r deparse(exact_formula)`\

### Covariate balance after matching
Note: The red dots are the balance against the entire pool of control candidates 
```{r love_plot, echo=F, message=FALSE, warning=FALSE}

love_df = df3 %>%
  filter(exposure==0) %>% 
  anti_join(matched_df %>% filter(exposure==0), by='guid_tz') %>% 
  bind_rows(matched_df )

if (params$psm == 1) {
  covariates = love_df %>%
  select(age_group,bmi_cat , simplified_vacc1_ww, SES, sex, sector_cat,sw_imunosuppressed, psw_preg)
} else {
  covariates = love_df %>%
  select(age, bmi ,vacc1_ww , pc_ses_comb20 , sex ,sector_cat,sw_imunosuppressed ,  psw_preg)
}


# data frame interface:
balanced_table = bal.tab(covariates,
                 treat = "exposure", data = love_df, un=T,
                 subclass = "subclass", disp.subclass = F,
                 s.d.denom = "pooled")


variables_mappings = list(sex ~ "Gender",
                   age ~ "Age",
                   age_group ~ "Age Group",
                   SES ~ "Socioeconomic status",
                   pn_flu_vacc_5yr ~ "Flu vaccination in the last 5 years",
                   bmi_cat ~ "BMI",
                   pc_smoke_desc ~ "Smoking status",
                   psw_cdc_certain_cancer ~ "Cancer (n,%)",
                   psw_cdc_certain_ckd ~ "Chronic kidney disease (n,%)",
                   psw_cdc_certain_copd ~ "Chronic obstructive pulmonary disease (n,%)",
                   psw_cdc_certain_heart ~ "Heart disease (n,%)",
                   overweight ~ "Overweight: BMI, 25 to 30",
                   obesity ~ "Obesity: BMI, 30 to 40",
                   severe_obesity ~ "Severe obesity: BMI 40+",
                   psw_preg ~ "Pregnancy",
                   psw_cdc_certain_sickle ~ "Sickle cell disease (n,%)",
                   psw_cdc_certain_dm2 ~ "Type 2 diabetes mellitus (n,%)",
                   psw_cdc_possible_asthma ~ "Asthma (n,%)",
                   psw_cdc_possible_cereb_vasc ~ "Cerebrovascular disease (n,%)",
                   psw_cdc_possible_other_resp ~ "Other respiratory disease (n,%)",
                   psw_cdc_possible_htn ~ "Hypertension (n,%)",
                   sw_imunosuppressed ~ "Immunosuppression (n,%)",
                   psw_cdc_certain_immuno ~ "Organ Transplant (n,%)",
                   psw_cdc_possible_neuro ~ "Neurological disease (n,%)",
                   psw_hepatic ~ "Liver disease (n,%)",
                   psw_dm_type1 ~ "Type 1 diabetes mellitus (n,%)",
                   sw_waned_immunity ~ "Waned Immunity", 
                   simplified_vacc1_ww ~ "first vaccine dose (weeks from campaign launch) ",
                   sector_cat ~ "Sector",
                   num_vacc_doses ~ "Num vaccine doses"
                   )



variables_mappings_vec = lapply(variables_mappings, rhs)
names(variables_mappings_vec)  =  lapply(variables_mappings, lhs)
variables_mappings_vec = append(variables_mappings_vec, c(age = "Age", bmi = "BMI", pc_ses_comb20 = "SES (20 levels)",  vacc1_ww = "first vaccine dose (weeks from campaign launch)",pn_pack_years = "Smoking Pack Years"))


```
```{r}
p = love.plot(balanced_table,
          drop.missing=F,
          disp.subclass=F,
          stats = c("mean.diffs"),
          sample.names = c("Unmatched", "Matched"),
          abs = T,
          var.names = variables_mappings_vec,
          line = F, 
          thresholds = c(m = .1))

    ggsave(glue("Results/covariate_balance.png"), print(p))
p
```

# per variable densities after matching
```{r}
plot_var(matched_df, pn_pack_years)
plot_var(matched_df, bmi)
plot_var(matched_df, age)
plot_var(matched_df, vacc1_ww_raw)
plot_var(matched_df, pc_ses_comb20)
```

```{r include=FALSE}
show_worst_matches(matched_df, bmi, simplified_bmi )
show_worst_matches(matched_df, age, age, quantile = 0.80, nrows = 20 )
show_worst_matches(matched_df, pn_pack_years, simplified_pack_years )
show_worst_matches(matched_df, pc_ses_comb20, simplified_ses )

```
```{r include=FALSE}
table_vars = c('exposure_cat','age', 'age_group', 'sector_cat', 'sex', 'SES', 'pn_flu_vacc_5yr', 'bmi_cat', 'pc_smoke_desc',  'psw_cdc_certain_cancer', 'psw_cdc_certain_ckd', 'psw_cdc_certain_copd', 'psw_cdc_certain_heart', 'severe_obesity', 'obesity', 'overweight', 'psw_preg', 'psw_cdc_certain_sickle', 'psw_cdc_certain_dm2', 'psw_dm_type1', 'psw_cdc_possible_asthma', 'psw_cdc_possible_cereb_vasc', 'psw_cdc_possible_other_resp', 'psw_cdc_possible_htn', 'sw_imunosuppressed', 'psw_cdc_certain_immuno', 'psw_cdc_possible_neuro', 'psw_hepatic', 'num_vacc_doses', 'sw_waned_immunity', 'simplified_vacc1_ww')

```


# Table3
#### matched population stratified by epxosure
```{r table3, echo=FALSE, message=FALSE, warning=FALSE}
if (params$create_table3  == 1) {
  table3 =  tbl_summary(
      matched_df ,
      by = exposure_cat, # split table by group
      label = variables_mappings,
      statistic = all_continuous() ~ "{mean} ({sd}), {median} ({p25},{p75})",
      missing = "ifany", # don't list missing data separately,
      include = all_of(table_vars)
      
    ) %>%
      modify_header(label = "**Variable**") %>% # update the column header
      modify_header(update = all_stat_cols() ~ "**{ifelse(level==0,'No Regeneron','Regeneron')}**, N = {n} ({style_percent(p, symbol = TRUE)})") %>%
      bold_labels() %>%
      add_overall() 
  table3 %>%
    gtsummary::as_flex_table() %>% 
    save_as_docx(path= glue("Results/table3.docx"))
  
  table3
  
}
```


# Table3.a
#### UNmatched population
```{r table3a, echo=FALSE, message=FALSE, warning=FALSE}
if (params$create_table3  == 1) {
  table3a =  tbl_summary(
      unmatched_df ,
      by = is_matched, 
      label = variables_mappings,
      statistic = all_continuous() ~ "{mean} ({sd}), {median} ({p25},{p75})",
      missing = "ifany", # don't list missing data separately,
      include = all_of(table_vars[table_vars != 'exposure_cat'])
      
    ) %>%
      modify_header(label = "**Variable**") %>% # update the column header
      modify_header(update = all_stat_cols() ~ "**{ifelse(level=='No','Unmatched Controls','Matched Controls')}**, N = {n} ({style_percent(p, symbol = TRUE)})") %>%  
      bold_labels() %>% 
      add_overall() 
  table3a %>%
    gtsummary::as_flex_table() %>% 
    save_as_docx(path= glue("Results/table3a.docx"))
  
  table3a
  
}
```


```{r include=FALSE}
filename = glue("feathers/matched/regeneron_effectivness_matched.feather")
write_feather(matched_df, filename)
```

