---
title: Survival via (Semi) Paramteric modeling
output: 
  html_document:
    toc: false
    toc_depth: 3
    toc_float: true
    code_folding: hide

date: "`r format(Sys.time(), '%d %B %Y')`"
params:
  sub_grp_anlsys: full
  psm: 0
 
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(comment='',
 root.dir = "/regeneron_effectiveness/code/R/",
  output_dir = "./Results",
  output_file = "Results/full_analysis.html",
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```


```{r message=FALSE, warning=FALSE, include=FALSE}

library(ggplot2)
library(glue)
library(dplyr)
library(feather)
library(cobalt)
library(knitr)
library(lubridate)
library(jtools)
library(flextable)
library(collections)
library(survminer)
library(car)
library(VIM) # for missing data
library(tidyverse)
```

```{r echo=FALSE}
source("./survival.R")
options(scipen = 999)

inner_params = dict(list(RD_factor = 1000.0,
                         perc = 2,
                         store_results=T,
                         fwp_days=30,
                         n_cores = 15,
                         type='prod',
                         fwp_days = 28,
                         n_iterations = 1000))


```

```{r}
input_table = "feathers/matched/regeneron_effectivness_matched.feather"
```

```{r}
df_raw = read_feather (input_table) 
```


## Analysis type: `r params$sub_grp_anlsys` \
### data file : `r input_table`\
### maximal followp days :  `r inner_params$get("fwp_days")`\
### administrative end of followup :  `r max(as.Date((df_raw$retrieval_date)))`\
### distance method for matching  (0 = mahalanobis, 1 = PS) =  `r params$psm` \


```{r echo=FALSE}


```{r}
df_severe = calc_survival_variables(df_raw, "first_time_severe_clean", params$sub_grp_anlsys,inner_params)
df_hosp   = calc_survival_variables(df_raw, "first_time_hosp", params$sub_grp_anlsys, inner_params)
df_death  = calc_survival_variables(df_raw, "date_ptira_corona", params$sub_grp_anlsys, inner_params)
```

```{r}
stratified_count = function(df) {
df %>% 
  group_by(exposure_cat, is_event) %>% 
  count() %>% 
    kable()
}

```
### hospitalization events count
```{r}
stratified_count(df_hosp)
```


### severe events count
```{r}
stratified_count(df_severe)
```

### death events count
```{r}
stratified_count(df_death)
```


### gender counts in current analysis : 
```{r}
df_hosp %>%  select(sex, exposure_cat) %>% group_by(sex) %>%  count() %>% kable()
```

```{r eval=FALSE, include=FALSE}

df_hosp %>% 
  group_by(psw_preg , simplified_vacc1_ww , sw_waned_immunity) %>% 
  count()
```
## verify No missing values
```{r echo=FALSE}
# Missing data plot
aggr(df_hosp, col=c('navyblue','red'), numbers=T, sortVars=TRUE, plot = F, labels=names(data), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))

```

```{r}
formula_str = 'Surv(days_of_fwp, is_event) ~ exposure_cat +  age  + as.factor(pn_flu_vacc_5yr) + bmi_cat + pc_smoke_desc   +   sex + as.factor(SES) + sector_cat + sw_imunosuppressed +   sw_waned_immunity + sw_heart + sw_diabetes+ sw_respiratory + psw_cdc_possible_htn + as.factor(simplified_vacc1_ww)+   psw_cdc_certain_ckd+ psw_cdc_certain_cancer+ psw_cdc_possible_neuro+ psw_hepatic'
```

## A glimpse on multi collinearity with num_vacc_doses
```{r}
car::vif(glm(is_event ~ exposure_cat +  age  + as.factor(pn_flu_vacc_5yr) + bmi_cat + pc_smoke_desc +  num_vacc_doses+  as.factor(SES) + sector_cat + sw_imunosuppressed +   sw_waned_immunity + sw_heart + sw_diabetes+ sw_respiratory + psw_cdc_possible_htn + as.factor(simplified_vacc1_ww)+   psw_cdc_certain_ckd+ psw_cdc_certain_cancer+ psw_cdc_possible_neuro+ psw_hepatic ,data = df_hosp, singular.ok = TRUE))
```

```{r}
car::vif(glm(is_event ~ exposure_cat +  age  + as.factor(pn_flu_vacc_5yr) + bmi_cat + pc_smoke_desc +  as.factor(SES) + sector_cat + sw_imunosuppressed +   sw_waned_immunity + sw_heart + sw_diabetes+ sw_respiratory + psw_cdc_possible_htn + as.factor(simplified_vacc1_ww)+   psw_cdc_certain_ckd+ psw_cdc_certain_cancer+ psw_cdc_possible_neuro+ psw_hepatic ,data = df_hosp, singular.ok = TRUE))
```

# Unadjusted Cox Model: hospitalizations
```{r}
get_cox_output(df_hosp,'Surv(days_of_fwp, is_event) ~ exposure_cat') 
```


# Adjusted Cox Model: hospitalizations
```{r}
get_cox_output(df_hosp,formula_str, filename= glue("cox_hospitalizations_{params$sub_grp_anlsys}.docx")) 
```

# Unadjusted Cox Model: Severe Covid-19
```{r echo=FALSE, message=FALSE, warning=FALSE}
get_cox_output(df_severe,'Surv(days_of_fwp, is_event) ~ exposure_cat') 
```

# Cox Model:Severe Covid-19
```{r echo=FALSE, message=FALSE, warning=FALSE}
get_cox_output(df_severe, formula_str, filename = glue("cox_severe_{params$sub_grp_anlsys}.docx")) 
```

# Unadjusted Cox Model: Death
```{r echo=FALSE, message=FALSE, warning=FALSE}
get_cox_output(df_death,'Surv(days_of_fwp, is_event) ~ exposure_cat') 
```

# Cox Model: Death 
```{r echo=FALSE}
 if (params$sub_grp_anlsys != 'below_60') {
  get_cox_output(df_death, formula_str, ph_test = F , filename = glue("cox_death_{params$sub_grp_anlsys}.docx"))  
 }
```


